Responsible Machine Learning Techniques\\\large{Interpretable Models, Post-hoc Explanation, and Discrimination Testing}

### 2019-12-19 article-information-2019.tex
nicks most recent content

\begin{itemize}[leftmargin=*,labelsep=5.8mm]
	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
	\item \textbf{Discrimination testing methods}: AIR, ME, and SMD -- \textsection\ref{ssec:di}
	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
\end{itemize}

@@

Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a template workflow for future users of similar methods and tools. Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discussed discrimination testing methods include AIR, ME, and SMD \cite{feldman2015certifying}, \cite{cohen2013statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.

@@

Metrics for testing discrimination in this text are:

@@

@@

\noindent where $\mathbf{X}_p$ and $\mathbf{X}_c$ represent binary markers created from some demographic attribute $\mathbf{X}_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and conditional probabilities, $\text{Pr}(\hat{\mathbf{y}}|\mathbf{X}_j = \alpha)$, are evaluated over all $\mathbf{x}^{(i)} \in \{\mathbf{X}_j | x^{(i)}_j = \alpha\}$. Accuracy in each demographic is also considered as a measure of validity for a model and a demographic group.

@@

\begin{table}[H]
\caption{Group size and accuracy for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
\centering
\begin{tabular}{cccc}
\toprule
\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& Accuracy \\
\midrule
Protected 1	& 3057.0		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.764\end{tabular} \\	
\midrule			 
Control 1		& 16943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.751\end{tabular} \\	
\midrule
Protected 2	& 9916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.761\end{tabular} \\	
\midrule
Control 2		& 10084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.745\end{tabular} \\	
\bottomrule
\label{tab:sim_n_acc}
\end{tabular}
\end{table}

\begin{table}[H]
\caption{AIR, ME and SMD for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
\centering
\begin{tabular}{cccccc}
\toprule
\textbf{Model}		& \textbf{Protected Class}								& \textbf{Control Class}							& \textbf{AIR}							& \textbf{ME}								& \textbf{SMD} \\
\midrule
$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}\\
\midrule 
$g^\text{XNN}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.737 \\ 1.04\end{tabular}	& \begin{tabular}{@{}c@{}} 11.1\%\\ -1.7\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.240 \\ 0.551\end{tabular} \\													
\bottomrule
\label{tab:sim_disc}
\end{tabular}
\end{table}

@@

\begin{table}[H]
\caption{Group size and accuracy for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
\centering
\begin{tabular}{cccc}
\toprule
\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& Accuracy \\
\midrule
Black 		& 2608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} \\	
\midrule			 
White 		& 28361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} \\	
\midrule
Female 		& 8301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular} \\	
\midrule
Male 		& 13166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} \\	
\bottomrule
\label{tab:mort_n_acc}
\end{tabular}
\end{table}

\begin{table}[H]
\caption{AIR, ME and SMD for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
\centering
\begin{tabular}{cccccc}
\toprule
\textbf{Model}		& \textbf{Protected Class}								& \textbf{Control Class}							& \textbf{AIR}							& \textbf{ME}								& \textbf{SMD} \\
\midrule
$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular} \\
\midrule 
$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular} \\

@@

\subsection{Interpretability, Explainability, Appeal, and Compliance}

Interpretable model architectures and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions. As seen in Sections \ref{sec:m_and_m} and \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain. Post-hoc explanations generally help describe and summarize mechanisms and decisions, potentially yielding an even clearer understanding of ML models. Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}. Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}. Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention. ML models are often wrong.\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.

@@

\subsection{Intersectionality of Interpretability, Explainability, Discrimination, and Security in ML}

>>>>

\begin{itemize}[leftmargin=*,labelsep=5.8mm]
	\item \textbf{Notation}: spaces, datasets, \& models -- \textsection\ref{ssec:not}
	\item \textbf{Training data}: simulated data \& collected mortgage data -- \textsection\ref{ssec:sim_data}~and~\textsection\ref{ssec:mort_data}
	\item \textbf{ML models}: constrained, interpretable MGBM \& XNN models  -- \textsection\ref{ssec:mgbm}~and~\textsection\ref{ssec:xnn}
	\item \textbf{Post-hoc explanation techniques}: PD, ICE, \& Shapley values -- \textsection\ref{ssec:pd_ice}~and~\textsection\ref{ssec:shap}
	\item \textbf{Discrimination testing methods}: AIR, ME, SMD and confusion matrix metrics -- \textsection\ref{ssec:di}
	\item \textbf{Software resources}: GitHub repository associated with Sections \ref{sec:m_and_m} and \ref{sec:res} -- \textsection\ref{ssec:soft}
\end{itemize}

@@

Post-hoc explanation and discrimination testing techniques are applied to constrained, interpretable models trained on the mortgage data to provide a template workflow for future users of similar methods and tools. Presented explanation techniques include PD, ICE, and Shapley values \cite{esl}, \cite{ice_plots}, \cite{shapley}, \cite{tree_shap}. PD, ICE, and Shapley values provide direct, global, and local summaries and descriptions of constrained models without resorting to the use of intermediary and approximate surrogate models. Discussed discrimination testing methods include AIR, ME, SMD, and confusion matrix metrics \cite{feldman2015certifying}, \cite{cohen1988statistical}, \cite{cohen1992power}.\footnote{Part 1607 - Uniform Guidelines on Employee Selection Procedures (1978) \textsection 1607.4.} Accuracy and other confusion matrix metrics are also reported by demographic segment \cite{zafar2017fairness}. All outlined materials and methods are implemented in open source Python code, and are made available in the software resources associated delineated in Subsection \ref{ssec:soft}.

@@

Because many technical and academic discussions of fairness in ML have been inconclusive\footnote{See: \href{https://www.youtube.com/watch?v=jIXIuYdnyyk}{Tutorial: 21 Fairness Definitions and Their Politics}.}, this text will draw on regulatory and legal standards that have been used for years in industries like Financial Services. The discussed metrics are also representative of fair lending analyses and pair well with the mortgage data. Before discussing the techniques, it is important to explain and draw a distinction between the two major types of discrimination recognized in US legal and regulatory settings, disparate treatment (DT), and disparate impact (DI). DT (which is loosely referred to as \textit{intentional discrimination}) occurs most often in an algorithmic setting when a model explicitly uses protected class status (e.g., race, sex) as an input feature or uses a feature that is so similar to protected class status that it essentially proxies for class membership.  With some limited exceptions, the use of these factors in an algorithm is illegal under several statutes in the US.\textsuperscript{4} DI, colloquially known as \textit{unintentional discrimination}, occurs when some element of a decisioning process includes a \textit{facially neutral} factor (i.e, a reasonable and valid predictor of response) that results in a disproportionate share of a protected class receiving an unfavorable outcome.  In modeling, this is most typically driven by a statistically important feature that is distributed unevenly across classes, which causes more frequent unfavorable outcomes for the protected class.  However, other factors, such as hyperparameter or algorithm choice, can drive DI.  Crucially, legality hinges on whether changing the model, for example exchanging one feature for another or altering the hyperparameters of an algorithm, can lead to a similarly predictive model with lower disparate impact.

The analyses and metrics herein focus on several measures of disparate impact that are commonly used in US litigation and regulatory settings. One is known as marginal effects (ME). ME is simply the difference between the percent of the control group members receiving a favorable outcome and the percent of the protected class members receiving a favorable outcome.

\noindent where $\mathbf{X}_p$ and $\mathbf{X}_c$ represent binary markers created from some demographic attribute $\mathbf{X}_j$, $c$ denotes the control group (often whites or males), $p$ indicates a protected group, and conditional probabilities, $\text{Pr}(\hat{\mathbf{y}}|\mathbf{X}_j = \alpha)$, are evaluated over all $\mathbf{x}^{(i)} \in \{\mathbf{X}_j | x^{(i)}_j = \alpha\}$.  ME is a favored DI metric used by the CFPB, the primary agency charged with regulating fair lending laws at the largest US lending institutions and various other participants in the consumer financial market.\footnote{See: \href{https://files.consumerfinance.gov/f/201510_cfpb_supervisory-highlights.pdf}{Supervisory Highlights, Issue 9, Fall 2015}.} Another important measure of DI is AIR, more commonly known as a \textit{relative risk ratio} in settings outside of regulatory compliance.

@@

\noindent AIR is equal to the ratio of the proportion of the protected class that receives a favorable outcome divided by the percent of the control class that receives a favorable outcome.  Another long-standing measure of DI is SMD. SMD is typically used to assess disparities in continuous variables, such as income differences in employment analyses, or interest rate differences in lending. It originated from work on statistical power, and is more formally known as \textit{Cohen’s d}. The SMD is equal to the difference in the average class outcomes minus the control class outcome, divided by a measure of the standard deviation of the population.\footnote{There are several measures of the standard deviation of the score that are typically used: \textbf{1.} the standard deviation of the population, irrespective of protected class status, \textbf{2.} a standard deviation calculated only over the two groups being considered in a particular calculation, or \textbf{3.} a pooled standard deviation, using the standard deviations for each of the two groups.} Cohen defined values of this metric to have \textit{small}, \textit{medium}, and \textit{large} effect sizes if the values exceeded 0.2, 0.5, and 0.8, respectively.

@@

\noindent The numerator in the SMD is equivalent to marginal effects but adds the standard deviation divisor as a standardizing factor.  Because of this standardization factor, the SMD has the advantage that it allows for a comparison across different types of variables, such as inequity in mortgage closing fees or inequities in the interest rates given on certain loans.  In this, one may apply definitions in \citet{cohen1988statistical} of \textit{small}, \textit{medium}, and \textit{large} effect sizes, which represent a measure of \textit{practical significance}, which is described in detail below.  Finally, confusion matrix metrics and their ratios in demographic groups are also considered as measures of DI in section \ref{sec:res}.

A finding of \textit{practical significance} means the disparity found is not only statistically significant, but also passes beyond a chosen threshold that would constitute prima facia evidence of illegal discrimination.  Its use represents a recognition that any large dataset is likely to show statistically significant differences in outcomes by class, even if those differences are not meaningful.  It further recognizes that there are likely to be situations where differences in outcomes are beyond a model user’s ability to correct them without significantly degrading the quality of the model. Moreover, practical significance is also needed model by model builders and compliance personnel to determine whether a model should undergo remediation efforts before it is put into production. Unfortunately, guidelines for practical significance, i.e., the threshold at which any statistically significant disparity would be considered evidence of illegal discrimination, are not as frequently codified as the standards for statistical significance.  One exception, however, is in employment discrimination analyses, where the US Equal Employment Opportunity Commission (EEOC) has stated that if the AIR is below 0.80 and statistically significant, then this constitutes prima facia evidence of discrimination, which the model user must rebut in order for the disparate impact not to be considered illegal discrimination.\footnote{Importantly, the standard of 0.80 is not a law, but a rule of thumb for agencies tasked with enforcement of discrimination laws.  ``Adoption of Questions and Answers To Clarify and Provide a Common Interpretation of the Uniform Guidelines on Employee Selection Procedures,'' Federal Register, Volume 44, Number 43 (1979).} It is important to note that the 0.80 measure of practical significance, also known as the \textit{80\% rule} and the \textit{4/5ths rule}, is explicitly used in relation to AIR, and it is not clear that the use of this threshold is directly relevant to testing fairness for metrics other than the AIR.

The legal thresholds for determining statistical significance is clearer and more consistent than that for practical significance.  The first guidance in U.S. courts occurred in a case involving discrimination in jury selection, \textit{Castaneda v. Partida}.\footnote{\textit{Castaneda v. Partida}, 430 US 482 - Supreme Court (1977)}   Here, the U.S. Supreme Court wrote that, ``As a general rule for such large samples, if the difference between the expected value and the observed number is greater than two or three standard deviations, then the hypothesis that the jury drawing was random would be suspect to a social scientist.''  This ``two or three standard deviations'' test was then applied to employment discrimination in \textit{Hazelwood School Districts v. United States.}\footnote{\textit{Hazelwood School Dist. v. United States}, 433 U.S. 299 (1977)} Out of this, a 5\% two-sided test (z=1.96), or an equivalent 2.5\% one-sided test, has become a common standard for determining whether evidence of disparities is statistically significant.

@@

Tables \ref{tab:sim_n_acc} and \ref{tab:sim_disc} show the results of the disparity tests using the simulated data for two hypothetical sets of class-control groups.  Several measures of disparities are shown, with the SMD calculated using the probabilities from the models, and the false positive rates (FPRs), false negative rates (FNRs), their ratios, MEs, and AIR calculated using a binary outcome based on a cutoff of 0.6 (anyone with probabilities of 0.6 or greater receives the favorable outcome).\footnote{The decision of what cutoff should be used in production is typically based on the model’s use case, rather than one based solely on the statistical properties of the predictions themselves.  For example, a model developer at a bank might build a credit model where the F1 score is maximized at a delinquency probability cutoff of 0.15.  For purposes of evaluating the quality of the model, she may review confusion matrix statistics (accuracy, recall, precision, etc.) using cutoffs based on the maximum F1 score. But, because of its risk tolerance and other factors, the bank itself might be willing to lend to anyone with a delinquency probability of 0.18 or lower, which would mean that anyone who is scored at 0.18 or lower would receive an offer of credit.  Because disparity analyses are concerned with how people are affected by the way the model is used, it is essential that any confusion matrix-based metrics of disparity be calculated on the in-production classification decisions, rather than the cutoffs that are not related to what those affected by the model will experience.} Since $g^{\text{MGBM}}$ and $g^{\text{XNN}}$ assumes that a higher score is favorable (as might be the case if the model were predicting responses to marketing offers), one might want to consider the relative false negative rates as a measure of the class-control disparities. Table \ref{tab:sim_disc} shows that protected group 1 has higher relative false negative rates under $g^{\text{XNN}}$ (1.14 vs. 1.06).  However, the overall false negative rates were lower for  $g^{\text{XNN}}$ (0.383 vs. 0.401). This illustrates a danger in considering relative class-control metrics in isolation when comparing across models: despite the $g^\text{MGBM}$ appearing to be a relatively fairer model, more protected group 1 members experience negative outcomes using $g^{\text{MGBM}}$. This is because FNR accuracy improves for both the protected group 1 and control group 1, but members of control group 1 benefit more than those in protected group 1.  Of course, the choice of which model is truly fairer is a policy question.

\begin{table}[H]
\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.} 
\centering
\begin{tabular}{cccccc}
\toprule
\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy}  								& \textbf{FPR}									& \textbf{FNR} \\
\midrule
Protected 1	& 3,057		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.770 \\ 0.764\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.150 \\ 0.167\end{tabular} 	& \begin{tabular}{@{}c@{}} 0.401 \\ 0.383\end{tabular} \\	
\midrule			 
Control 1		& 16,943		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.739 \\ 0.751\end{tabular}	& \begin{tabular}{@{}c@{}} 0.129 \\ 0.149\end{tabular}	& \begin{tabular}{@{}c@{}} 0.378 \\ 0.337\end{tabular} \\
\midrule
Protected 2	& 9,916		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.758 \\ 0.761\end{tabular}	& \begin{tabular}{@{}c@{}} 0.169 \\ 0.183\end{tabular}	& \begin{tabular}{@{}c@{}} 0.331 \\ 0.306\end{tabular} \\	
\midrule
Control 2		& 10,084		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.729 \\ 0.745\end{tabular}	& \begin{tabular}{@{}c@{}} 0.091 \\ 0.116\end{tabular}	& \begin{tabular}{@{}c@{}} 0.420 \\ 0.370\end{tabular} \\		
\bottomrule
\label{tab:sim_n_acc}
\end{tabular}
\end{table}

11.1\% fewer control group 1 members receive the favorable offer under the ME column in Table \ref{tab:sim_disc}.  Of note is that 11.1\% is not a meaningful difference without context.  If the population of control group 1 and control group 2 were substantially similar in relevant characteristics, 11.1\% could represent an extremely large difference and would require remediation.  But if they represent substantially different populations, then 11.1\% could represent a reasonable deviation from parity.  As an example, if a lending institution that has traditionally focused on high credit quality clients were to expand into previously under-banked communities, an 11.1\% class-control difference in loan approval rates might be expected because the average credit quality of the new population would be lower than that of the existing population. Protected group 1’s AIR under $g^\text{XNN}$ is 0.737, is below the EEOC 4/5ths rule.  It is also highly statistically significant (not shown), which together would indicate that there may be evidence of illegal DI.  As with ME and other measures, the reasonableness of this disparity is not clear outside of context.  However, most regulated institutions that do perform discrimination analyses would find an AIR of this magnitude concerning and warranting further review.

\begin{table}[H]
\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the simulated test data.}
\centering
\begin{tabular}{cccccccc}
\toprule
\textbf{Model}		& \textbf{Protected Class}										& \textbf{Control Class}								& \textbf{AIR}									& \textbf{ME}									& \textbf{SMD}   								& \textbf{FPR Ratio}							& \textbf{FNR Ratio} \\
\midrule
$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.752 \\ 1.10\end{tabular}	& \begin{tabular}{@{}c@{}} 9.7\% \\ -3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.206 \\ 0.106\end{tabular}	& \begin{tabular}{@{}c@{}} 1.16 \\ 1.86\end{tabular}	& \begin{tabular}{@{}c@{}} 1.06 \\ 0.788\end{tabular}\\
\midrule 
$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Protected 1 \\ Protected 2\end{tabular}		& \begin{tabular}{@{}c@{}} Control 1 \\ Control 2\end{tabular}	& \begin{tabular}{@{}c@{}} 0.737 \\ 1.04\end{tabular}	& \begin{tabular}{@{}c@{}} 11.1\%\\ -1.7\%\end{tabular}	& \begin{tabular}{@{}c@{}} -0.240 \\ 0.551\end{tabular}	& \begin{tabular}{@{}c@{}} 1.12 \\ 1.59\end{tabular}	& \begin{tabular}{@{}c@{}} 1.14 \\ 0.827\end{tabular} \\										
\bottomrule
\label{tab:sim_disc}
\end{tabular}
\end{table}

\noindent Even though SMD is often used to measure disparities for non-categorical outcomes, it is calculated here on the models’ probabilities prior to being transformed into classifications.  This measurement would be particularly relevant if the probabilities are being used in combination with other models to determine an outcome.  The results show that $g^\text{MGBM}$ has less disparate impact than $g^\text{XNN}$ (-0.21 versus -0.24), but both are close to Cohen’s small effect threshold of -0.20.   Whether a small effect would be a highlighted concern would depend on a organization’s chosen threshold for flagging models for further review.

@@

\begin{table}[H]
\caption{Group size, accuracy, FPR, and FNR for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
\centering
\begin{tabular}{ccccccc}
\toprule
\textbf{Class}	& $\mathbf{N}$	& \textbf{Model} 												& \textbf{Accuracy} 									& \textbf{FPR}											& \textbf{FNR} \\
\midrule
Black 		& 2,608		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.654 \\ 0.702\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.315 \\ 0.295\end{tabular} 			& \begin{tabular}{@{}c@{}} 0.457 \\ 0.308\end{tabular} \\	
\midrule			 
White 		& 28,361		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.817 \\ 0.857\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.150 \\ 0.120\end{tabular}			& \begin{tabular}{@{}c@{}} 0.508 \\ 0.360\end{tabular}\\		
\midrule
Female 		& 8,301		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}}  0.768 \\ 0.822\end{tabular}		& \begin{tabular}{@{}c@{}} 0.208 \\ 0.158\end{tabular}			& \begin{tabular}{@{}c@{}} 0.402 \\ 0.322\end{tabular}\\
\midrule
Male 		& 13,166		& \begin{tabular}{@{}c@{}} $g^\text{MGBM}$ \\ $g^\text{XNN}$\end{tabular}	& \begin{tabular}{@{}c@{}} 0.785 \\	0.847\end{tabular} 		& \begin{tabular}{@{}c@{}} 0.182 \\ 0.131\end{tabular}			& \begin{tabular}{@{}c@{}} 0.497 \\ 0.347\end{tabular}\\
\bottomrule
\label{tab:mort_n_acc}
\end{tabular}
\end{table}

\begin{table}[H]
\caption{AIR, ME, SMD, FPR ratio, and FNR ratio for $g^\text{MGBM}$ and $g^\text{XNN}$ on the mortgage test data.}
\centering
\begin{tabular}{cccccccc}
\toprule
\textbf{Model}	& \textbf{Protected Class}	& \textbf{Control Class}	& \textbf{AIR}	& \textbf{ME}	& \textbf{SMD}  & \textbf{FPR Ratio}		& \textbf{FNR Ratio} \\
\midrule
$g^\text{MGBM}$	& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.776 \\ 0.948\end{tabular}	& \begin{tabular}{@{}c@{}} 18.3\% \\ 4.1\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.628 \\ 0.084\end{tabular}	& \begin{tabular}{@{}c@{}} 2.10 \\ 1.15\end{tabular}	& \begin{tabular}{@{}c@{}} 0.900 \\ 0.810\end{tabular} \\
\midrule 
$g^\text{XNN}$		& \begin{tabular}{@{}c@{}} Black \\ Female\end{tabular}		& \begin{tabular}{@{}c@{}} White \\ Male\end{tabular}	& \begin{tabular}{@{}c@{}} 0.743 \\ 0.955\end{tabular}	& \begin{tabular}{@{}c@{}} 21.4\% \\ 3.6\%\end{tabular}	& \begin{tabular}{@{}c@{}} 0.621 \\ 0.105\end{tabular}	& \begin{tabular}{@{}c@{}} 2.45 \\ 1.21\end{tabular}	& \begin{tabular}{@{}c@{}} 0.855 \\ 0.927\end{tabular} \\

@@

\subsection{Appeal and Override of Automated Decisions}

Interpretable model architectures and post-hoc explanations play an important role in increasing transparency into model mechanisms and predictions. As seen in Sections \ref{sec:m_and_m} and \ref{sec:res}, interpretable models often enable users to enforce domain knowledge-based constraints on model behavior, to ensure that models obey reasonable expectations, and to gain data-derived insights into the modeled problem domain. Post-hoc explanations generally help describe and summarize mechanisms and decisions, potentially yielding an even clearer understanding of ML models. Together they can allow for human learning from ML, certain types of regulatory compliance, and crucially, human appeal or override of automated model decisions \cite{art_and_sci}. Interpretable models and post-hoc explanations are likely good candidates for ML uses cases under the FCRA, ECOA, GDPR and other regulations that may require explanations of model decisions, and they are already used in the financial services industry today for model validation and other purposes.\footnote{See: \href{https://ww2.amstat.org/meetings/jsm/2019/onlineprogram/AbstractDetails.cfm?abstractid=303053}{\textit{Deep Insights into Explainability and Interpretability of Machine Learning Algorithms and Applications to Risk Management}}.}\textsuperscript{,}\footnote{Unfortunately, many non-consistent explanation methods can result in drastically different global and local feature importance values across different models trained on the same data or even for refreshing a model with augmented training data \cite{molnar}. Consistency and accuracy guarantees are perhaps a factor in the growing momentum behind Shapley values as a candidate technique for generating consumer-specific adverse action notices for explaining and appealing automated ML-based decisions in highly-regulated settings such as credit lending \cite{bracke2019machine}.} Writ large, transparency in ML also facilitates additional responsible AI processes such as model debugging, model documentation, and logical appeal and override processes, some which may also be required by applicable regulations.\footnote{E.g.: \href{https://www.federalreserve.gov/supervisionreg/srletters/sr1107.htm}{US Federal Reserve Bank Supervision and Regulation (SR) Letter 11-7: Guidance on Model Risk Management}.} Among these, appeal may deserve the most attention. ML models are often wrong\footnote{``All models are wrong, but some are useful.'' -- George Box, Statistician (1919 - 2013)} and appealing black-box decisions is difficult.\textsuperscript{2} For high-stakes, human-centered, or regulated applications that are trusted with mission- or life-critical decisions, the ability to logically appeal or override inevitable wrong decisions is not only a possible prerequisite for regulatory compliance, but also an important failsafe procedure for those affected by ML decisions.

@@

\subsection{Intersectional and Non-static Problems in Machine Learning}
